Simple repression draft

### 2020-05-29 doc/appendix_03_bayesian_inference.tex
MRM pass through MM section on inference for unreg promoters for SI

The different models explored in this work embraced different levels of 
coarse-graining that resulted in a diverse set of number of parameters. For each
of these model configurations Bayes theorem demands from us to represent our 
preconceptions on the possible parameter values in the form of the prior 
$P(\vec{\theta})$. Throughout this work for models with $> 1$ parameter we 
assign independent priors to each of the parameters; this is
\begin{equation}
P(\vec{\theta}) = \prod_{i=1}^n P(\theta_i).
\end{equation}
Although it is common practice to use non-informative, or maximally
uninformative priors, we are of the mindset that this is a disservice to the
philosophical and practical implications of Bayes theorem. It sounds almost
contradictory to claim that can we represent our thinking about a natural
phenomena in the form of a mathematical model -- which itself is a bold claim of
deeply understanding the features that matter in the object of study -- but we
have absolutely no idea what the parameter values ought to be. We therefore make
use of our own expertise, many times in the form of order-of-magnitude
estimates, to write down prior distributions for our parameters.

@@

To proceed we need to specify a prior. In this case we are extremely data-rich,
as the dataset from Jones et.\ al~\cite{Jones2014} has of order 1000-3000
single-cell measurements for each promoter, so our choice of prior matters
little here, as long as it is sufficiently broad. A convenient choice for our
problem is to use a \textit{conjugate} prior. A conjugate prior is a special
prior that causes the posterior to have the same functional form as the prior,
simply with updated model parameters. This makes calculations analytically
tractable and also offers a nice interpretation of the inference procedure as
updating our knowledge about the model parameters. This makes conjugate priors
very useful when they exist. The caveat is that conjugate priors only exist for
a very limited number of likelihoods, mostly with only one or two model
parameters, so in almost all other Bayesian inference problems, we must tackle
the posterior numerically.

@@

Since we did use a conjugate prior, we may simply look up our posterior in any
standard reference such as~\cite{Gelman2013}, Section 2.6,
from which we find that
\begin{equation}
\lambda
\sim \text{Gamma}\left(\alpha + \bar{m}N, \beta + N\right),
\end{equation}
where we defined the sample mean $\bar{m} = \frac{1}{N}\sum_k m_k$ for
notational convenience. A glance at the FISH data from~\cite{Jones2014} reveals
that $N$ is $\mathcal{O}(10^3)$ and $\langle m\rangle \gtrsim 0.1$ for all
constitutive strains in~\cite{Jones2014}, so $\bar{m}N \gtrsim 10^2$. Therefore
as we suspected, our prior parameters are completely overwhelmed by the data.
The prior behaves, in a sense, like $\beta$ extra ``data points''
with a mean value of $(\alpha-1)/\beta$~\cite{Gelman2013}, which
gives us some intuition for how much data is needed to overwhelm
the prior in this case: enough data $N$ such that $\beta\ll N$
and $\alpha/\beta \ll \bar{m}$. In
fact, $\bar{m}N$ and $N$ are so large that we can, to an excellent
approximation, ignore the $\alpha$ and $\beta$ dependence and approximate the
gamma distribution as a Gaussian with mean $\bar{m}$ and standard deviation
$\sqrt{\bar{m}/N}$, giving
\begin{equation}
\lambda
\sim \text{Gamma}\left(\alpha + \bar{m}N, \beta + N\right)
\approx \text{Normal}\left(\bar{m}, \sqrt{\frac{\bar{m}}{N}}\right).
\end{equation}
As an example with real numbers, for the \textit{lacUV5} promoter, Jones et.\
al~\cite{Jones2014} measured 2648 cells with an average mRNA count per cell of
$\bar{m} \approx 18.7$. In this case then, our posterior is
\begin{equation}
\lambda
\sim \text{Normal}\left(18.7, 0.08\right),
\end{equation}
which suggests we have inferred our model's one parameter to a precision of
order 1\%.

This is not wrong, but it is not the full story. The model's posterior is
tightly constrained, but is it a good generative model? In other words, does the
model generate data that look similar to our actual data, and is it therefore
plausible that the model captures the important features of the data generating
process? This intuitive notion can be codified with \textit{posterior predictive
checks}, or PPCs, and we will see that this simple Poisson model fails badly.

The intuitive idea of posterior predictive checks is simple: \mmnote{Not sure if
all this stuff is clearer by example, but then does it confuse the reader in
generalizing to other models?}
\begin{enumerate}
\item Make a random draw of the model parameter $\lambda$ from the posterior
distribution.
\item Plug that draw into the likelihood and generate a synthetic dataset
$\{m_k\}$ conditioned on $\lambda$.
\item Repeat many times.
\end{enumerate}
More formally, the posterior predictive distribution can be thought of as the
distribution of future yet-to-be-observed data, conditioned on the data we have
already observed. Clearly if those data appear quite different, the model has a
problem. Put another way, if we suppose the generative model is true, then the
synthetic datasets we generate should resemble the actual observed data, and if
not, it suggests the model is missing important features. All the data we
consider in this work are 1D (distributions of mRNA counts over a population) so
ECDFs are an excellent visual means of comparing synthetic and observed
datasets. In general for higher dimensional datasets, much of the challenge is
in merely designing good visualizations that can actually show if synthetic and
observed data are similar or not.

For our example Poisson promoter model then, we merely draw many random numbers,
say 1000, from the Gaussian posterior. For each one of those draws, we generate
a dataset from the likelihood, i.e., we draw 2648 (the number of observed cells
in the actual dataset) Poisson-distributed numbers for each of the 1000
posterior draws, for a total of 2648000 samples from the posterior predictive
distribution.

To compare so many samples with the actual observed data, one excellent
visualization for 1D data is ECDFs of the quantiles, as shown for our Poisson
model in~\fig{fig:constit_post_full}.
In this example, the median for each possible mRNA
count is shown as a dark green line, while shaded bands of lighter
green contain 95\% of the posterior predictive samples.
Plotting quantiles in this way gives us a sense of the range of data we might
consider plausible, under the assumption that the model is true. In this case it
is quite obvious that the observed data, plotted in orange, could not plausibly
come from this Poisson generative model. The other model with a PPC plotted
in~\fig{fig:constit_post_full}, bursty model 5 with
a negative binomial steady-state mRNA
distribution, looks like a good candidate by eye. We cover it in more detail
later.

\subsubsection{Model 5 - Bursty promoter}
Let us now consider the problem of parameter inference from FISH data for model
five from~\fig{fig1:means_cartoons}(C). As derived in
Appendix~\ref{sec:gen_fcn_appdx}, the steady-state mRNA distribution in this
model is a negative binomial distribution, given by
\begin{equation}
p(m) = \frac{\Gamma(m+k_i)}{\Gamma(m+1)\Gamma(k_i)}
        \left(\frac{1}{1+b}\right)^{k_i}
        \left(\frac{b}{1+b}\right)^m,
\label{eq:neg_bionom}
\end{equation}
where $b$ is the mean burst size and $k_i$ is the burst rate nondimensionalized
by the mRNA degradation rate $\gamma$. As sketched earlier, the story of this
distribution is of geometrically-distributed bursts of mRNA, where the arrival
of bursts is a Poisson process with rate $k_i$ and the mean size of a burst is
$b$.

As for the Poisson promoter model, this expression for the steady-state mRNA
distribution is exactly the likelihood we want to use in Bayes theorem. Again
denoting the single-cell mRNA count data as $D=\{m_1, m_2,\dots, m_N\}$, here
Bayes theorem takes the form
\begin{equation}
p(k_i, b \mid D) \propto p(D\mid k_i,b)p(k_i, b).
\end{equation}
We already have our likelihood --the product of $N$ negative binomials as
Eq.~\ref{eq:neg_bionom}--  so we only need to choose priors on $k_i$ and $b$.
For the datasets from~\cite{Jones2014} that we are analyzing, as for the Poisson
promoter model above we are still data-rich so the prior's influence remains
weak, but not nearly as weak because the dimensionality of our model has
increased from one to two.

@@

Consider $k_i$. The fastest known bacterial promoters control
rRNA genes and initiate transcripts no faster than $\sim
1/\text{sec}$. It would be exceedingly strange if any of the
constitutive promoters from~\cite{Jones2014} were stronger than
that, so we can take that as an upper bound. For a lower bound,
if transcripts are produced too rarely, there would be nothing to
see with FISH. The datasets for each strain contain of order
$10^3$ cells, and if the $\langle m \rangle = k_i b/\gamma
\lesssim 10^{-2}$, then the total number of expected mRNA
detections would be single-digits or less and we would have
essentially no data on which to carry out inference. So assuming
$b$ is not too different from 1, justified next, and an mRNA
lifetime of $\gamma^{-1}\sim 3-5~\text{min}$, this gives us soft
bounds on $k_i/\gamma$ of perhaps $10^{-2}$ and $3\times 10^1$.

@@

With this, we are ready to write our full generative model as
\begin{equation}
\begin{split}
\ln k_i \sim \text{Normal}(-0.5, 2)
\\
\ln b \sim \text{Normal}(0.5, 1)
\\
m \sim \text{NBinom}(k_i, b).
\end{split}
\end{equation}
We carried out MCMC sampling on the posterior of this model, starting
with the constitutive \textit{lacUV5} dataset from~\cite{Jones2014}.
The resulting MCMC samples are shown in~\fig{fig:constit_post_full}(A).
In constrast to the active/inactive constitutive model considered
in~\cite{Razo-Mejia2020} (nonequilibrium model four
in~\ref{fig2:constit_cartoons}), this model is well-identified
with both parameters determined to a fractional uncertainty of 5-10\%.
The strong correlation reflects the fact that their product sets
the mean of the mRNA distribution, which is tightly constrained
by the data, but there is weak ``sloppiness''~\cite{Transtrum2015}
along a set of values with a similar product.

Flush with success having found the model's posterior to be
well-identified, the next step was posterior predictive sampling.
The only procedural difference compared to the Poisson promoter
model is that, since we have no analytical posterior expression
but only samples from the posterior, we generate a synthetic
dataset (i.e., posterior predictive samples) for each posterior
sample from the MCMC sampling. \fig{fig:constit_post_full}(B)
shows the resulting predictive ECDF; similarly to the Poisson
promoter model, the median posterior predictive ECDF is plotted
as a dark blue line, while a lighter blue shaded region
encompasses 95\% of the posterior predictive samples. Unlike the
Poisson promoter model, the experimental ECDF closely tracks the
posterior predictive ECDF, indicating this model is actually able
to generate the observed data and increasing our confidence that
this model is at least not wrong.

\textit{lacUV5} is our primary target here, since it forms the
core of all the simple repression constructs of~\cite{Jones2014}
that we consider in Section~\ref{sec:rep_kinetics_inference}.
Nevertheless, we thought it wise to apply our bursty promoter
model to the other 17 constitutive promoters available in the
FISH dataset from~\cite{Jones2014} as a test that the model is
capturing the essential phenomenology. If the model fit well to
all the different promoters, this would increase our confidence
that it would serve well as a foundation for inferring repressor
kinetics later in Section~\ref{sec:rep_kinetics_inference}.
Conversely, were the model to fail on more than a couple of the
other promoters, it would give us pause.

\fig{fig:constit_post_full}(C) shows the results, plotting the
posterior distribution from individually MCMC sampling all 18
constitutive promoter datasets from~\cite{Jones2014}. To aid
visualization, rather than plotting samples for each promoter's
posterior as we did in \fig{fig:constit_post_full}(A), for each
posterior we find and plot the curve that surrounds the 95\%
highest probability density region, i.e., on a plot analogous
to~(A), each contour would enclose approximately 95\% of the
samples, and thus 95\% of the probability mass,
of its posterior distribution.
Posterior predictive ECDFs (not shown) display a similar level of
agreement between data and predictive samples as for the bursty
model with \textit{lacUV5} in \fig{fig:constit_post_full}(B).
\mmnote{Might be worth showing some/all of these PPCs in
supplement for completeness? We could do postage stamp sized diff
plots and probably fit all 18 promoters on 1-2 pages.}

One interesting feature from \fig{fig:constit_post_full}(C) is
that burst rate varies far more widely, over a range of
$\sim10^2$, than burst size, confined to a range of
$\lesssim10^1$ (and with the exception of promoter 6, just a span
of 3-5x). This suggests that $k_i$, not $b$, is the key dynamic
variable that promoter binding site tunes.

It is interesting to connect this observation to the work
of~\cite{Brewster2012}, where these same 18 promoters were
considered through the lens of the three-state equilibrium model
and binding energies $\Delta\epsilon_P$ were predicted from an
energy matrix model derived from~\cite{Kinney2010}. The connection
is made simply by equating mean mRNA levels in both theories,
\mmnote{derivation is hasty, needs elaboration}
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma}
        \frac{\frac{P}{N_{NS}}\exp(-\beta\Delta\epsilon_P)}
                {1+\frac{P}{N_{NS}}\exp(-\beta\Delta\epsilon_P)},
\end{equation}
and taking the weak-promoter limit in the equilibrium model,
\begin{equation}
\langle m \rangle = \frac{k_i b}{\gamma}
        = \frac{r}{\gamma} \frac{P}{N_{NS}}\exp(-\beta\Delta\epsilon_P),
\end{equation}
valid for all the binding energies considered here.

How are the two coarse-grainings related? The only association that makes dimensional sense and produces the correct order-of-magnitude for the known parameters is to take
\begin{equation}
\frac{k_i}{\gamma} = \exp(-\beta\Delta\epsilon_P)
\label{eq:bursty_equil_corresp1_appdx}
\end{equation}
and
\begin{equation}
b = \frac{r}{\gamma}\frac{P}{N_{NS}}.
\label{eq:bursty_equil_corresp2_appdx}
\end{equation}
\fig{fig:constit_post_full}(D) shows that this linear scaling
between $\ln k_i$ and $-\beta\Delta\epsilon_P$ is approximately
true for all 18 constitutive promoters considered. The intercept
is an uninteresting function of the unknown model parameters, so
we only consider the slope of the scaling in
\fig{fig:constit_post_full}(D).
\mmnote{Wait, is that true or can we do something with the intercept, related to the burst size in some useful way? Needs more careful thought.}

While the associations represented by
\eq{eq:bursty_equil_corresp1_appdx} and \eq{eq:bursty_equil_corresp2_appdx}
appear to be born out by the scaling $\ln k_i \sim
-\beta\Delta\epsilon_P$, we do not find the association of
parameters encoded by
\eqrange{eq:bursty_equil_corresp1_appdx}{eq:bursty_equil_corresp2_appdx}
intuitive. If some deeper insight is waiting to be gleaned from it,
we leave it as an open problem.

\mmnote{Outline, still to cover:
\begin{itemize}
\item All 18 promoters. Hey the model still works on things other than UV5,
cool. Interesting and somewhat counterintuitive scaling b/w burst rate and
binding energy. Puzzle in comparing w/ Chong2014 supercoiling model. Are we
missing important things?? Unclear, we leave as open question. Good enough for
our purposes: posterior is identifiable, PPC is great, and of the models we've
thought of it is unique in satisfying both.
\end{itemize}
}

\mmnote{Discuss puzzling comparison w/ Chong2014: if supercoiling is the thing, why are my burst sizes all the same but burst rates vary? And why is the duty cycle of my promoter so low, ie., bursts so short? COmpare their fig 7E; their $\beta/\alpha$ is my $k^+/k^-$. They have one or two genes with very small $\beta/\alpha$, does the \textit{galK} locus just happen to be that, or is there a deeper disagreement? Hard to say w/o more data.}

\subsection{Transcription factor kinetics can be inferred from FISH}

>>>>

The different models explored in this work embraced different levels of
coarse-graining that resulted in a diverse number of parameters for different
models. For each of these model configurations Bayes theorem demands from us to
represent our preconceptions on the possible parameter values in the form of the
prior $P(\vec{\theta})$. Throughout this work for models with $> 1$ parameter we
assign independent priors to each of the parameters; this is
\begin{equation}
P(\vec{\theta}) = \prod_{i=1}^n P(\theta_i).
\end{equation}
Although it is common practice to use non-informative, or maximally
uninformative priors, we are of the mindset that this is a disservice to the
philosophical and practical implications of Bayes theorem. It sounds almost
contradictory to claim that can we represent our thinking about a natural
phenomena in the form of a mathematical model -- which itself is a bold claim of
deeply understanding the features that matter in the object of study -- but we
have absolutely no idea what the parameter values ought to be. We therefore make
use of our own expertise, many times in the form of order-of-magnitude
estimates, to write down prior distributions for our parameters. In summary then
we don’t want to be maximally uninformative because we aren’t completely
uninformed.

\mrm{Still need to write something about the intuition of being data-rich and 
how then the prior matters little.}

@@

Having introduced the ideas behind Bayesian inference we are ready to apply the
theoretical machinery to our non-equilibrium models. In particular in this
section we will focus on model 1 and model 5 in
Figure~\ref{fig2:constit_cartoons}(A). Model 1, the Poisson promoter, will help
us build practical intuition into the implementation of the Bayesian inference
pipeline as we noted in Section~\ref{sec:beyond_means} of the main text that
this model cannot be reconciled with experimental data from observables such as
the Fano factor. In other words, we acknowledge that this model is ``wrong,''
but we still see value in going through the analysis since the simple nature of
the model translates into a neat statistical analysis.

\subsubsection{Model 1 - Poisson promoter}

As specified in the main test, the mRNA steady-state distribution for model 1 in
Figure~\ref{fig2:constit_cartoons}(A) is Poisson with parameter $\lambda$.
Throughout this Appendix we will appeal to the convenient notation for
probability distributions of the form
\begin{equation}
m \sim \text{Poisson}(\lambda),
\end{equation}
where the simbol ``$\sim$'' can be read as \textit{is distributed according to}.
So the previous equation can be read as: the mRNA copy number $m$ is distributed
according to a Poisson distribution with parameter $\lambda$. Our objective then
is to compute the posterior probability distribution $P(\lambda \mid D)$, where,
as in the main text, $D = \{ m_1, m_2, \ldots, m_N \}$ are the data consisting
of single-cell mRNA counts. Since we can assume that each of the cells mRNA
counts are independent of any other cells, our likelihood function $P(D \mid
\lambda)$ consists of the product of $N$ Poisson distributions.

To proceed with the inference problem we need to specify a prior. In this case
we are extremely data-rich, as the dataset from Jones et.\ al~\cite{Jones2014}
has of order 1000-3000 single-cell measurements for each promoter, so our choice
of prior matters little here, as long as it is sufficiently broad. A convenient
choice for our problem is to use a \textit{conjugate} prior. A conjugate prior
is a special prior that causes the posterior to have the same functional form as
the prior, simply with updated model parameters. This makes calculations
analytically tractable and also offers a nice interpretation of the inference
procedure as updating our knowledge about the model parameters. This makes
conjugate priors very useful when they exist. The caveat is that conjugate
priors only exist for a very limited number of likelihoods, mostly with only one
or two model parameters, so in almost all other Bayesian inference problems, we
must tackle the posterior numerically.

@@

Since we did use a conjugate prior, we may simply look up our posterior in any
standard reference such as~\cite{Gelman2013}, Section 2.6,
from which we find that
\begin{equation}
\lambda
\sim \text{Gamma}\left(\alpha + \bar{m}N, \beta + N\right),
\end{equation}
where we defined the sample mean $\bar{m} = \frac{1}{N}\sum_k m_k$ for
notational convenience. A glance at the FISH data from~\cite{Jones2014} reveals
that $N$ is $\mathcal{O}(10^3)$ and $\langle m\rangle \gtrsim 0.1$ for all
constitutive strains in~\cite{Jones2014}, so $\bar{m}N \gtrsim 10^2$. Therefore
as we suspected, our prior parameters are completely overwhelmed by the data.
The prior behaves, in a sense, like $\beta$ extra ``data points''
with a mean value of $(\alpha-1)/\beta$~\cite{Gelman2013}, which
gives us some intuition for how much data is needed to overwhelm
the prior in this case: enough data $N$ such that $\beta\ll N$
and $\alpha/\beta \ll \bar{m}$. In
fact, $\bar{m}N$ and $N$ are so large that we can, to an excellent
approximation, ignore the $\alpha$ and $\beta$ dependence and approximate the
gamma distribution as a Gaussian with mean $\bar{m}$ and standard deviation
$\sqrt{\bar{m}/N}$, giving
\begin{equation}
\lambda
\sim \text{Gamma}\left(\alpha + \bar{m}N, \beta + N\right)
\approx \text{Normal}\left(\bar{m}, \sqrt{\frac{\bar{m}}{N}}\right).
\end{equation}
As an example with real numbers, for the \textit{lacUV5} promoter, Jones et.\
al~\cite{Jones2014} measured 2648 cells with an average mRNA count per cell of
$\bar{m} \approx 18.7$. In this case then, our posterior is
\begin{equation}
\lambda
\sim \text{Normal}\left(18.7, 0.08\right),
\label{eq:gauss_posterior}
\end{equation}
which suggests we have inferred our model's one parameter to a precision of
order 1\%.

This is not wrong, but it is not the full story. The model's posterior
distribution is tightly constrained, but is it a good generative model? In other
words, if we use the model to generate synthetic data in the computer does it
generate data that look similar to our actual data, and is it therefore
plausible that the model captures the important features of the data generating
process? This intuitive notion can be codified with \textit{posterior predictive
checks}, or PPCs, and we will see that this simple Poisson model fails badly.

The intuitive idea of posterior predictive checks is simple: 
\begin{enumerate}
\item Make a random draw of the model parameter $\lambda$ from the posterior
distribution.
\item Plug that draw into the likelihood and generate a synthetic dataset
$\{m_k\}$ conditioned on $\lambda$.
\item Repeat many times.
\end{enumerate}
More formally, the posterior predictive distribution can be thought of as the
distribution of future yet-to-be-observed data, conditioned on the data we have
already observed. Clearly if those data appear quite different, the model has a
problem. Put another way, if we suppose the generative model is true, i.e. we
claim that our model explains the process through which our observed
experimental data was generated, then the synthetic datasets we generate should
resemble the actual observed data. If this is not the case, it suggests the
model is missing important features. All the data we consider in this work are
1D (distributions of mRNA counts over a population) so empirical cumulative
distribution functions ECDFs are an excellent visual means of comparing
synthetic and observed datasets. In general for higher dimensional datasets,
much of the challenge is in merely designing good visualizations that can
actually show if synthetic and observed data are similar or not.

For our example Poisson promoter model then, we merely draw many random numbers,
say 1000, from the Gaussian posterior in Eq.~\ref{eq:gauss_posterior}. For each
one of those draws, we generate a dataset from the likelihood, i.e., we draw
2648 (the number of observed cells in the actual dataset) Poisson-distributed
numbers for each of the 1000 posterior draws, for a total of 2648000 samples
from the posterior predictive distribution.

To compare so many samples with the actual observed data, one excellent
visualization for 1D data is ECDFs of the quantiles, as shown for our Poisson
model in~\fig{fig:constit_post_full}(B) in the main text.

\subsubsection{Model 5 - Bursty promoter}

Let us now consider the problem of parameter inference from FISH data for model
five from~\fig{fig1:means_cartoons}(C). As derived in
Appendix~\ref{sec:gen_fcn_appdx}, the steady-state mRNA distribution in this
model is a negative binomial distribution, given by
\begin{equation}
p(m) = \frac{\Gamma(m+k_i)}{\Gamma(m+1)\Gamma(k_i)}
        \left(\frac{1}{1+b}\right)^{k_i}
        \left(\frac{b}{1+b}\right)^m,
\label{eq:si_neg_bionom}
\end{equation}
where $b$ is the mean burst size and $k_i$ is the burst rate nondimensionalized
by the mRNA degradation rate $\gamma$. As sketched earlier, we can intuitively
think about this distribution through a simple story. The story of this
distribution is that the promoter undergoes geometrically-distributed bursts of
mRNA, where the arrival of bursts is a Poisson process with rate $k_i$ and the
mean size of a burst is $b$.

As for the Poisson promoter model, this expression for the steady-state mRNA
distribution is exactly the likelihood we want to use in Bayes theorem. Again
denoting the single-cell mRNA count data as $D=\{m_1, m_2,\dots, m_N\}$, here
Bayes theorem takes the form
\begin{equation}
p(k_i, b \mid D) \propto p(D\mid k_i,b)p(k_i, b),
\end{equation}
where the likelihood $p(D\mid k_i,b)$ is given by the product of $N$ negative
binomials as in Eq.~\ref{eq:si_neg_bionom}. We only need to choose priors on
$k_i$ and $b$. For the datasets from~\cite{Jones2014} that we are analyzing, as
for the Poisson promoter model above we are still data-rich so the prior's
influence remains weak, but not nearly as weak because the dimensionality of our
model has increased from one to two.

@@

Consider $k_i$. Some of the strongest known bacterial promoters control rRNA
genes and initiate transcripts no faster than $\sim 1/\text{sec}$. It would be
exceedingly strange if any of the constitutive promoters from~\cite{Jones2014}
were stronger than that, so we can take that as an upper bound. For a lower
bound, if transcripts are produced too rarely, there would be nothing to see
with FISH. The datasets for each strain contain of order $10^3$ cells, and if
the $\langle m \rangle = k_i b/\gamma \lesssim 10^{-2}$, then the total number
of expected mRNA detections would be single-digits or less and we would have
essentially no data on which to carry out inference. So assuming $b$ is not too
different from 1, justified next, and an mRNA lifetime of $\gamma^{-1}\sim
3-5~\text{min}$, this gives us soft bounds on $k_i/\gamma$ of perhaps $10^{-2}$
and $3\times 10^1$.

@@

With this, we are ready to write our full generative model as
\begin{equation}
\begin{split}
\ln k_i \sim \text{Normal}(-0.5, 2),
\\
\ln b \sim \text{Normal}(0.5, 1),
\\
m \sim \text{NBinom}(k_i, b).
\end{split}
\end{equation}
Section ~\ref{section_04_bayesian_inference} in the main text details the
results of applying this inference to the single-cell mRNA counts data.

\subsection{Bayesian inference on the simple-repression architecture}
